---
title: "Phyloseq_PreProcessing"
author: "Yining Sun"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: united
    highlight: tango
editor_options: 
  chunk_output_type: console
---

# Prepare to start the preprocessing
```{r start-prep}
#Be in the right place
setwd("/local/workdir/ys636/BIOMI6300-Project2")

# load packages
pacman::p_load(tidyverse, phyloseq, ggplot2, install = FALSE)

# source functions
source("/local/workdir/ys636/BIOMI6300-Project2/code/functions.R")
source("/local/workdir/ys636/BIOMI6300-Project2/code/colors_and_shapes.R")

```


# Load the data
```{r load-data}

# load in the raw_physeq data from DADA2 Practice.Rmd
load("/local/workdir/ys636/BIOMI6300-Project2/data/raw_physeq.RData")

# Take a look at the S4 object
raw_phyloseq

# Tell us about your phyloseq
str(raw_phyloseq)
typeof(raw_phyloseq)

View(raw_phyloseq@otu_table)
View(raw_phyloseq@tax_table)
View(data.frame(sample_data(raw_phyloseq)))

```


# Remove unnecessary taxa
```{r rm-ASVs}

# 1. Mitochondria ASVs
# 2. Chloroplast ASVs
# Note: since I don't have controls or mock community, I will skip removing this two types of ASV data.

test_physeq <-
  raw_phyloseq %>% 
   ## remove the mitochondria ASVs
  subset_taxa(Family != "Mitochondria") %>%
   ## remove the chloroplast ASVs
  subset_taxa(Order != "Chloroplast")

noMitoChloroCJMock_physeq <- 
  raw_phyloseq %>%
  subset_taxa(Family != "Mitochondria" | is.na(Family)) %>%
  subset_taxa(Order != "Chloroplast" | is.na(Order)) %>%
  prune_samples((sample_names(.)) %!in% c("CJ-V08-P", "MockZyomPos"), .) %>%
  # now we will remove any ASVs of count 0
  prune_taxa(taxa_sums(.) > 0, .)
  

noMitoChloroCJMock_physeq


# how many taxa have we removed so far?
num_ASVs_rm <- ntaxa(raw_phyloseq) - ntaxa(noMitoChloroCJMock_physeq) ## 0
# proportion
prop_ASVs_rm <- ntaxa(raw_phyloseq)/ntaxa(noMitoChloroCJMock_physeq)  ## 1

### After this step, there is no taxa removed,
### which means there is no mitochondria or chloroplast ASVs in the samples

```


# Evaluate sequence depth
```{r seq-depth}

# calculate the sum of the seqs in each sample
seqSums_df <- 
  noMitoChloroCJMock_physeq %>%
  otu_table() %>%
  # take the sum of each column of the ASV matrix
  colSums() %>%
  data.frame() %>%
  rownames_to_column(var = "names") 
  
# Add "source" column from metadata_df to seqSums_df
seqSums_df <- cbind(seqSums_df, metadata[!names(metadata) %in% names(seqSums_df)])

head(seqSums_df)


# fix column name in 2nd column
colnames(seqSums_df)[2] <- "Totalseqs"

head(seqSums_df)


# plot it
# histogram plot
seqSums_df %>%
  ggplot(aes(x = Totalseqs)) +
  geom_histogram(bins = 50) +
  labs(title = "Histogram of raw seq counts per sample")

### From this histogram plot, we can see 12 samples only have less than 10,000 seqs, 
### and the rest 4 samples are kind of progressing arithmetically.



# density plot
seqSums_df %>%
  ggplot(aes(x = Totalseqs, fill = source)) +
  geom_density(alpha = 0.5) 

### From the density plot, we can see dairy group samples have more seqs than meat group samples,
### in general.


```

### I decided to proceed though neither the histogram nor the density plot 
### looks fair enough for later analysis.
### I think it is worth trying!

### I would not move any samples since:
### 1. I don't have many samples;
### 2. the first 12 samples shown in the histogram have pretty similar seq depth.


# Save preprocessed phyloseq
```{r}
save(noMitoChloroCJMock_physeq, file = "/local/workdir/ys636/BIOMI6300-Project2/data/preprocessed_physeq.RData")

### I can also directly use raw_physeq.RData since after this workflow, no ASVs are removed.
```


